---
title: "Meeting 3 Analysis Script"
author: "Daniel Conroy-Beam"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(reshape2)
```



```{r load}

#Load in the data
data<-read.csv("Meeting 2/Meeting 2 Data PROCESSED.csv")

#Recode variable
data$PIN<-as.factor(data$PIN)
data$gender<-as.factor(data$gender)
data$shockCause<-as.factor(data$shockCause)
data$shock<-as.factor(data$shock)
data$moral<-as.factor(data$moral)

```



## Descriptives

Descriptive statistics for guilt:
```{r descriptivesGuilt}

c(summary(data$guilt),"SD"=sd(data$guilt,na.rm=T))

```


Descriptive statistics for generosity:
```{r descriptivesGenerosity}

#Pre-shock:
c(summary(data$preShare),"SD"=sd(data$preShare,na.rm=T))

#Post-shock:
c(summary(data$postShare),"SD"=sd(data$postShare,na.rm=T))

```



Descriptive statistics for personality:
```{r descriptivesPersonality}

#Agreeableness
c(summary(data$agree),"SD"=sd(data$agree,na.rm=T))

#Neuroticism
c(summary(data$neuro),"SD"=sd(data$neuro,na.rm=T))

```



## Guilt Analysis

### Prediction 1.1

We expected people would feel more guilt when they caused the shock vs. their partner.

This prediction was supported. People felt more guilt when they caused a shock to their partner.

```{r p1.1}

#Run the t-test for prediction 1.1
p1.1<-t.test(data$guilt ~ data$shockCause)

#Output the result
p1.1
```



### Prediction 1.2

We expected that people would feel more guilt when their partner was morally good vs. morally bad.

This prediction was supported. People felt more guilt when their partner was morally good.
```{r p1.2}

#Run the t-test for prediction 1.2
p1.2<-t.test(data$guilt~data$moral)

#Output the result
p1.2
```



### Prediction 1.3

We expected participants would report more guilt when their partner received a strong shock.

This prediction was supported. Guilt increased with increasing shock magnitude.
```{r p1.3}

#Run the ANOVA for prediction 1.3
p1.3<-lm(guilt~shock,data=data)

#Output the result
anova(p1.3)

#Output the group means
tapply(data$guilt, data$shock, function(x) mean(x,na.rm=T))

```



### Prediction 2

We expected that women will report more guilt than men

This prediction was not supported. There was no significant difference in guilt between men and women.
```{r p2}

#Run the t-test
p2<-t.test(data$guilt~data$gender)

#Output the result
p2
```



### Prediction 3
We expected agreeable people would report more guilt.

This prediction was supported. Agreeableness was weakly but significantly positively correlated with reported levels of guilt.
```{r p3}

#Run the correlation test for prediction 3
p3<-cor.test(data$agree,data$guilt)

#Output the result
p3

```



### Prediction 4
We expected neurotic people would report more guilt.

This prediction was not supported. There was no relationship between neuroticism and guilt
```{r p4}

#Run the correlation test
p4<-cor.test(data$neuro,data$guilt)

#Output the result
p4

#Test predictions 3 and 4 simultaneously using a regression
guiltReg<-lm(guilt~neuro + agree, data=data)

#Output regression result
summary(guiltReg)
```




## Generosity

```{r melt}

#Melt the data for within-subjects analysis
dataMelt<-melt(data,
               measure.vars=c("preShare","postShare"),
               variable.name=c("prePost"),
               value.name=c("share"))

```



### Prediction 5.1

We expected that people would be more generous when they caused the shock vs. when the partner caused the shock.

This prediction was supported. Participants were more generous to their partner after the shock when the participant thought they themselves (rather than the partner) caused the shock.
```{r p5.1}

#Run the within-subjectes ANOVA for prediction 5.1
p5.1<-aov(share ~ shockCause*prePost + Error(PIN), data=dataMelt)

#Output the ANOVA results:
summary(p5.1)

#Output the means separated by condition
tapply(dataMelt$share,dataMelt[,c("shockCause","prePost")],
       function(x) mean(x,na.rm=T))
```



### Prediction 5.2

We expected participants would be more generous when the partner was morally good vs. morally bad.

This prediction was supported. Post-shock, participants were more generous toward morally good partners.

```{r p5.2}

#Run the within-subjects ANOVA
p5.2<-aov(share~prePost*moral+Error(PIN),data=dataMelt)

#Output the results
summary(p5.2)

#Output means as a function of condition
tapply(dataMelt$share,
       dataMelt[,c("moral","prePost")],
       function(x) mean(x,na.rm=T))

```



### Prediction 5.3

We expected participants would be more generous when the shock was stronger.

This prediction was supported. Amount of generosity after the shock increases with shock magnitude.
```{r p5.3}

#Run the within-subjects ANOVA
p5.3<-aov(share~prePost*shock+Error(PIN),data=dataMelt)

#Output the results
summary(p5.3)

#Output means as a function of condition
tapply(dataMelt$share,
       dataMelt[,c("shock","prePost")],
       function(x) mean(x,na.rm=T))

```



### Prediction 6

We predicted that women would be more generous than men.

This prediction was not supported. There was not a significant interaction between gender and generosity.

```{r p6}

#Run the within-subjects ANOVA
p6<-aov(share~prePost*gender+Error(PIN),data=dataMelt)

#Output the results
summary(p6)

#Output means as a function of condition
tapply(dataMelt$share,
       dataMelt[,c("gender","prePost")],
       function(x) mean(x,na.rm=T))

```



### Prediction 7

We expected that agreeable people would be more generous than disagreeable people.

This prediction was partially supported. Agreeable people were more generous after the shock but not before.

```{r p7}

#Run correlation test for pre-shock generosity
p7pre<-cor.test(data$preShare,data$agree)

#Output the results
p7pre

#Run the correlation test for post-shock generosity
p7post<-cor.test(data$postShare,data$agree)

#Output the results
p7post

```



### Prediction 8

We expected that neurotic people would be more generous than emotionally stable people.

This prediction was not supported. There is no correlation between neuroticism and generosity either before or after the shock.

```{r p8}

#Run correlation test for pre-shock generosity
p8pre<-cor.test(data$preShare,data$neuro)

#Output the results
p8pre

#Run the correlation test for post-shock generosity
p8post<-cor.test(data$postShare,data$neuro)

#Output the results
p8post

```